Formal Analysis of EDHOC Key Establishment for Constrained IoT Devices

### 2022-07-14 paper.tex
Changed through Section 3.2

%-------------------------------------------------------------------------- sec
\section{\uppercase{Denial of Service}}
\label{sec:dos}

% There is little about EDHOC specific DoS stuff; that is interesting in itself,
% and motivates this section
There is a wealth of information about DoS in general, and more specifically for
IoT.
%
However, there is little information regarding DoS aspects stemming
specifically from \mEdhoc{}.
%

% Other protocols have DoS considerations, but EDHOC delegates it to developer
In contrast to \mTls{} and IKEv2,
the \mEdhoc{} core specification does not include any specific measures for
countering Denial of Service (DoS) attacks.
%
Instead, that responsibility is delegated to the application, using
proprietary measures or extensions to \mEdhoc{}.
%

% The core is small and probably not vulnerable to anything unexpected
The \mEdhoc{} core specification has limited functionality, few negotiable
parameters and processing rules.
%
Hence, we argue in this section that its attack surface
is limited from a DoS perspective.
%

% However, It is difficult to define what EDHOC *is*; it s partially specified in many
% specs
However, \mEdhoc{} leaves much of the semantics of its messages undefined or
specified as extensions.
%
Examples of mechanisms with mostly undefined semantics include processing
rules and content of error messages and of the additional data that is
passed in each of the \mEdhoc{} messages.
%
Another example is that the core specification does not include a
message transport mechanism.
%
The transport for \mEdhoc's primary target, i.e., to key \mOscore{},
is \mCoap{}.
%
This is specified in~\cite{ietf-core-oscore-edhoc-03}, but still leaves
many design choices to the application.
%
Because of this, it is difficult to draw the line between what is considered as
\mEdhoc{} and what is an extension.
%

% Pros with undefined semantics: flexibilty; Cons: developers can shoot
% themselves in the foot
Undefined semantics allow the implementors much freedom
and flexibility to design applications.
%
On the down-side, the few recommendations given increase the risk of
applications being vulnerable to DoS via \mEdhoc{}, even
though the core \mEdhoc{} specification itself is arguably not particularly
weak in this respect.
%

@@

No recommendation is given on how to minimize DoS effects stemming from use of
error messages, which can be used in an arbitrary way in \mEdhoc{}
(see Section~\ref{sec:errorHandling} below).
%
While there is some work on DDoS aspects of \mCoap{}, and much on IoT in
general, this does not capture the particular consequences originating with
\mEdhoc{}.
%
We now elaborate on some considerations application developers needs to take
into account.
%

@@

Depending on the use case, the first \mEdhoc{} message may be sent by either
party in a session.
%
For example, even though it may seem intuitive that an electronic
key-card initiates the connection towards lock, the first \mEdhoc{} message
could be included in the response from the lock.
%
This means the lock can postpone heavy verification actions until it has
authenticated the key card after receiving the third \mEdhoc{} message.
%
The draft IETF specification profiling \mEdhoc{} for use with
\mCoap{} and \mOscore{} ~\cite{ietf-core-oscore-edhoc-03},
however, defines the
\mCoap{} client to be the \mEdhoc{} initiator, and hence reduces the flexibility
to use that option.
%

@@

\textbf{Q:} Is there any situation where it makes sense that the initiator sends
a unsupported ciphersuite message instead of message 3?
%

@@

Section 6.2 gives an example of sending an error message "method not
supported".
%
This message cannot be used to negotiate method securely.
%
The reason is that error messages are not integrity protected, and an initator
only proposes one method, and therefore the same tactic used for ciphersuite
negotiation cannot be used for methods.
%
Because method negotiation can be viewed in the context of crypto agility, there
is room for improvement.
%

>>>>

%-------------------------------------------------------------------------- sec
\section{\uppercase{Denial of Service}}
\label{sec:dos}
%\mPoint{DoS is out of scope for \mEdhoc{}; apps to deal with it}
\mEdhoc{} includes no specific measures for countering Denial of Service (DoS)
attacks and delegates the responsibility to handle this to the application.
%
The core \mEdhoc{} specification has limited functionality, few negotiable
parameters and processing rules.
%
Hence, we argue in this section that its attack surface
is limited from a DoS perspective.
%
However, \mEdhoc{} leaves much of the semantics of its messages undefined or
specified as (optional) extensions.
%
It is therefore difficult to draw the line between what is \mEdhoc{} and what is
an extension.
%
For example, the core specification does not include a
message transport mechanism.
%
This increases the risk of applications vulnerable to DoS via \mEdhoc{}, even
though \mEdhoc{} itself is arguably not particularly weak in this respect.
%

@@

No recommendation is given on how to minimize DoS effects stemming from use of
error messages, which can be used in an arbitrary way in \mEdhoc{}.
%
While there is some work on DDoS aspects of \mCoap{}, and much on IoT in
general, this does not capture the particular consequences originating with
\mEdhoc{}.
%
We now elaborate on some considerations application developers needs to take
into account.
%

@@

Depending on the use case, the first \mEdhoc{} message may be sent by either
party in a session.
%
For example, even though it may seem intuitive that an electronic
key-card initiates the connection towards lock, the first \mEdhoc{} message
could be included in the response from the lock.
%
This means the lock can postpone heavy verification actions until it has
authenticated the key card after receiving the third \mEdhoc{} message.
%

@@

\textbf{Q:} Is htere any situation where it makes sense that the initiator sends
a unsupported ciphersuite message instead of message 3?
%

@@

Section 6.2 gives an excample of sending an error message "method not
supported".
%
This message cannot be used to negotiate method securely.
%
The reason is that error messages are not integrity protected, and an initator
only proposes one method, and therefore the same tactic used for ciphersuite
negotiation cannot be used for methods.
%
Because method negotiation can be viewed in the context of crypto agility, there
is room for improvement.
%
